---
title: "Single-case (N-of-1) inference with PhysioClinStats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single-case (N-of-1) inference with PhysioClinStats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

PhysioClinStats is the clinical inference engine of the ecosystem: mixed-effects
and MMRM longitudinal models, single-case (N-of-1) designs, estimands with
multiple imputation, and per-estimate uncertainty. The heavy modelling backends
(`lme4`, `mmrm`, `emmeans`, `mice`, ...) are optional, guarded `Suggests`.

This vignette uses only the **distribution-free single-case tools**, which need
no modelling backend, so it runs anywhere. Each estimator reproduces its
reference implementation (`SingleCaseES`, `scan`) exactly.

```{r setup}
library(PhysioClinStats)
```

## 1. A single case (AB design)

We have one participant measured repeatedly: a baseline phase (A) and an
intervention phase (B). Higher is better (e.g. comfortable gait speed, scaled).

```{r data}
A <- c(20, 20, 26, 25, 22)          # baseline
B <- c(28, 25, 30, 29, 31, 30)      # intervention
```

## 2. Non-overlap effect sizes

The non-overlap family summarises how cleanly the intervention points separate
from baseline. `resultValue()` pulls the numbers out of the returned
`AnalysisResult`.

```{r nonoverlap}
pnd <- scedPND(A, B)   # percentage of non-overlapping data
pem <- scedPEM(A, B)   # percentage exceeding the median
nap <- scedNAP(A, B)   # non-overlap of all pairs (= AUC), with a CI

PhysioExperiment::resultValue(pnd)$estimate
PhysioExperiment::resultValue(pem)$estimate
PhysioExperiment::resultValue(nap)[c("estimate", "ci_lower", "ci_upper", "p_value")]
```

Tau rescales NAP to [-1, 1]; Tau-U additionally corrects for a baseline trend
(and exposes both the `SingleCaseES` and `scan` denominator conventions).

```{r tau}
PhysioExperiment::resultValue(scedTau(A, B))$estimate
PhysioExperiment::resultValue(scedTauU(A, B, method = "parker"))$estimate
```

## 3. Two-SD band

A Shewhart-style band flags runs of intervention points beyond 2 SD of the
baseline mean -- a transparent visual-analysis aid.

```{r band}
band <- scedTwoSDBand(A, B)
res <- PhysioExperiment::resultValue(band)
res[c("upper", "lower", "first_run_at")]
```

## 4. Frame the question as an estimand

Even for a single case it helps to state the estimand explicitly (ICH E9(R1)):
the treatment, population, endpoint, intercurrent-event strategy and summary
measure.

```{r estimand}
defineEstimand(
  treatment = "task-oriented training",
  population = "one post-stroke participant",
  endpoint = "comfortable gait speed",
  intercurrent = list(event = "missed sessions",
                      strategy = "treatment-policy"),
  summary_measure = "non-overlap of all pairs"
)
```

## Where to go next

- Longitudinal group models -- `fitMixedModel()`, `fitMMRM()`,
  `estimatedMarginalMeans()` -- require their modelling backends and guard for
  them at run time.
- Estimands with missing data -- `multipleImputation()`, `poolEstimates()`,
  `analyseEstimand()`.
- Causal tools -- `causalMediation()`, `targetTrialEmulate()`.
- `conformalPredict()` gives distribution-free prediction intervals.

See `?PhysioClinStats` and each function's help page for details.
